---
title: "Week of 07-25-2022: Lit Review"
author: "Zachary Rowson"
date: "7/26/2022"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(readxl)
library(DT)
```

## Introduction

This document evaluates the concordance of gabi results with results from other studies.
  
This comparison will be useful for assessing the results of analysis in terms of other
laboratory results. Do chemicals found active tend to be active in other papers and do
chemicals inactive tend to be inactive? Will also help assess the effect that including
more endpoints has on detection of chemical activity. Are we seeing activity in chemicals
widely active in other studies or are we potentially seeing activity previously missed or spurious activity
using these endpoints. Also, identifying chemicals widely active in other studies, but not active
in the gabi analysis, may provide insight into the lack of assay sensitivity.

## Format Literature Review

```{r, data-load, message=FALSE, warning=FALSE}
load("~/Desktop/Stephanie Projects/DNT60 Analysis/pipelining/results/Padilla_DNT60_tcpl_out.Rdata")
tcplOut.dt <- as.data.table(DNT60_tcpl_out)

# Load literature review table and save sheet of interest as data.table

litReview <- as.data.table( readxl::read_xlsx("../tables/copy_Chemical Comparison Tables_Behavior Results and  Literature_FINAL-rev5_2_22-jko.xlsx",
                               sheet = "gabiComparison") )
```
```{r, data-format, warning=FALSE}
# Set column names and row names
newNames <- c( "cpid", sapply(names(litReview), gsub, pattern=" ", replacement=".")[-1] )
setnames(litReview, old = names(litReview), new = newNames)
litReview[1, cpid := "5,5-Diphenylhydantoin"]
litReview[12, cpid := "Bis(tributyltin) Oxide"]
litReview[52, cpid := "Polybrominated diphenyl ether (PBDE)-47"]

# ID chemicals active in net activity metrics
active.1 <- tcplOut.dt[endp%in%c("avgS_L","avgS_D","avgS_T") & hitcall>0.8, unique(name)]

# ID chemicals active in other endpoints
active.2 <- tcplOut.dt[!endp%in%c("avgS_L","avgS_D","avgS_T","AUC_L","AUC_D","AUC_T") & hitcall>0.8,
                       unique(name)]

# Edit litReview columns
litReview[is.na(Active.in.gabi.analysis), Active.in.gabi.analysis := 0]
litReview[cpid %in% active.1, active.in.Net.Activity.metrics := 1]
litReview[Active.in.gabi.analysis==1 & is.na(active.in.Net.Activity.metrics),
          active.in.Net.Activity.metrics := 0]
litReview[cpid %in% active.2, Active.in.other.endpoints := 1]
litReview[Active.in.gabi.analysis==1 & is.na(Active.in.other.endpoints),
          Active.in.other.endpoints := 0]

# str(litReview)
litReview[, Active.in.gabi.analysis := as.logical(Active.in.gabi.analysis) ]
litReview[, Number.of.publications.with.activity := as.integer(Number.of.publications.with.activity)]
litReview[, `%.of.papers.with.activity` := as.numeric(`%.of.papers.with.activity`)]

litReview[Number.of.asociated.publications==0,
          `:=` (Number.of.publications.with.activity=NA, `%.of.papers.with.activity`=NA)]

DT::datatable(litReview, caption="Formatted Literature Review table.",
              extensions = 'FixedColumns',
              options = list(scrollX=TRUE),
              rownames = FALSE)
```

## Questions and Answers

For the 21 chemicals found active, on average how many papers from other studies are reporting activity?

```{r, concordance-active}
litReview[Active.in.gabi.analysis==TRUE, mean(`%.of.papers.with.activity`,na.rm=TRUE)]
```

On average we are seeing that papers associated with active chemicals are 74.44% active.

Does this differ for inactive chemicals?

```{r, concordane-inactive}
litReview[Active.in.gabi.analysis==FALSE, mean(`%.of.papers.with.activity`,na.rm=TRUE)]
```

On average we are seeing that 56.00% of papers associated with a gabi inactive chemical report activity. So, papers associated with chemicals active in gabi analysis tend to have more associated papers that found activity.

Would it be better to just do some binomial like thing i.e. what is the probability that a paper associated with a gabi active chemical reports activity vs. a paper associated with a gabi inactive?

```{r, percent-active}
litReview[Active.in.gabi.analysis==TRUE, sum(Number.of.publications.with.activity,na.rm=T) / sum(Number.of.asociated.publications)]
litReview[Active.in.gabi.analysis==F, sum(Number.of.publications.with.activity,na.rm=T) / sum(Number.of.asociated.publications)]
```

We see similar numbers as above calculation for this value, 76.19% for actives and 55.36% for inactives. Using binomial distribution may make modeling and determining statistical significance easier, though draws are not i.i.d. .

Do the percent of chemicals without any associated papers differ from inactive to active chemicals.

```{r}
n_1 <- litReview[Active.in.gabi.analysis==TRUE & Number.of.asociated.publications==0, .N]
litReview[Active.in.gabi.analysis==TRUE, n_1 / .N]
n_2 <- litReview[Active.in.gabi.analysis==FALSE & Number.of.asociated.publications==0, .N]
litReview[Active.in.gabi.analysis==FALSE, n_2 / .N]
```

Active chemicals have no associated publications less often than inactive chemicals (32.82% v. 44.75%), though if this difference is statistically significant hasn't been established.

How does active concordance in studies differ between chemicals found active in the commonly used net activity metric, those found active in other endpoints, and those active in both categories?

```{r}
litReview[active.in.Net.Activity.metrics==T & Active.in.other.endpoints==F,
          mean(`%.of.papers.with.activity`,na.rm=TRUE)]
litReview[active.in.Net.Activity.metrics==F & Active.in.other.endpoints==T,
          mean(`%.of.papers.with.activity`,na.rm=TRUE)]
litReview[active.in.Net.Activity.metrics==T & Active.in.other.endpoints==T,
          mean(`%.of.papers.with.activity`,na.rm=TRUE)]
```

Chemicals active in net activity metrics but not other endpoints (only one chemical, BPA) were active in 83.33% of associated papers. Chemicals active in other endpoints were active in 68.75% of associated papers on average. Chemicals active in net activity and other endpoints were active in 80.56% of papers on average.

The difference in percent active could be a result of spurious detection of activity or could be an indication that evaluation of these other endpoints detects chemical perturbation missed by other laboratories.

To get at Katie's question of whether evaluation of data using these other endpoints results in detection of chemical activity where chemical effect may have been missed, look at chemicals active in other endpoints only, and see if any were tested in lots of other studies and found inactive.

```{r}
DT::datatable(litReview[active.in.Net.Activity.metrics==F & Active.in.other.endpoints==T,][order(-Number.of.asociated.publications), .(cpid, Number.of.asociated.publications, Number.of.publications.with.activity, `%.of.papers.with.activity`)],
              extensions = 'FixedColumns',
              options = list(scrollX=TRUE),
              rownames = FALSE)
```

There aren't any chemicals with many associated publications but a small percent that observed chemical effect. Diethylstilbesterol we see 2 total publications both with no activity. 

Running tangential to this, there are some chemicals that have no associated publications, which could be an indication that the chemical has been overlooked for some reason.

```{r}
litReview[active.in.Net.Activity.metrics==F & Active.in.other.endpoints==T &
            Number.of.asociated.publications==0, cpid]
```

Chemicals active in other endpoints but without any publications were 5,5-Diphenylhydantoin, Heptachlor epoxide, and Triethyltin. May be that Heptachlor epoxide is assumed to be too teratogenic? Triethyltin as well? 5,5-Diphenylhydantoin is interesting. Obviously there may be publications that tested these chemicals that were not found.

Were there any gabi inactive chemicals that were widely tested and widely found active in those other studies?

```{r}
litReview[Active.in.gabi.analysis==F & Number.of.asociated.publications>=3, .(cpid, Number.of.asociated.publications, Number.of.publications.with.activity, `%.of.papers.with.activity`)][order(-Number.of.asociated.publications)]
```

Valproate is the most extreme example where it was found active in 6 of 7 associated publications. (I think Valproate was a near miss in some endpoints). Could the lack of activity in gabi be due to it's teratogenic properties? (Isn't is teratogenic?) This is probably oversimplifying, but could Valproate be used as an example of what this assay can be used for, the screening of chemicals for behavioral effects in the absence of morphological effect. Valproate is known to be teratogenic in humans but also neurodevelopmentally toxic. The utility of this assay may be identifiying behavioral effects in the absence of moprphological effect which is likely more rare. May help explain the lack of sensitivity.

We also see activity in many publications for lead acetate and acetaminophen. Is the same argument possible?

Inactive chemicals with 3 associated publications with 2 or 3 papers with activity are Haloperidol, Nicotine, and Carbamezipine.
